timesliced_random_cell_selection

### 2018-11-07 src/raster_sjl.Rmd
lat lon on background points done (+ little progress on the raster brick loop)

# load the netcdf and raster packages
```{r}
library(ncdf4)
library(raster)
```

@@

Because I am using these files for Maxent, I need the projection to be an albers equal area. We can project the whole raster brick. There are a couple of ways to do this, but the simplest way is to use an existing file in the projection I want. The file I will use for this is called aea.tif. I've sent it to you in the email with this code as a zip. Note that there are two files in the zip folder - you will need both (but you only call on the .tif in the code) 
```{r}
aea <- raster("../output/env/aea.tif") # change the file path to where ever the tif and .twf files are saved
rasbr_aea <- projectRaster(rasbr, aea) 
plot(rasbr_aea[[1]]) # just to look at the top layer of the raster brick after projection so you can see the difference
```

@@

#the loop....
netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
library (ncdf4)
for (i in 1:no_netcdf){  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working on
  brkyr <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2) # extracting the second part of the netcdf filename (which is the month)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:length(obssp)) {  
      print(j)
      de <- obssp$depth[[j]]  # a variable for the observation depth layer
      yr <- obssp$year[j]  # a variable for the observation year
      mth <- obssp$month[j]  # a variable for the observation month
      #I think i need an if statement here... if brkyr == yr & brkmth == mth do the data extraction below
        if (brkyr == yr & brkmth == mth){
          obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp) 
          obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)

        }
    }
}
obssp$salinity_depth
obssp$salinity_surface
```
not extracting the correct values...

@@

>>>>

# load the netcdf and raster packages
```{r}
library(ncdf4)
library(raster)
library(rgdal)
```

@@

Because I am using these files for Maxent, I need the projection to be an albers equal area. We can project the whole raster brick. There are a couple of ways to do this, but the simplest way is to use an existing file in the projection I want. The file I will use for this is called aea.tif. I've sent it to you in the email with this code as a zip. Note that there are two files in the zip folder - you will need both (but you only call on the .tif in the code) 
```{r}
aea <- raster("../output/env/aea.tif") # change the file path to where ever the tif and .twf files are saved
rasbr_aea <- projectRaster(rasbr, aea) 
plot(rasbr_aea[[1]]) # just to look at the top layer of the raster brick after projection so you can see the difference
rasbr_aea
```

@@

#the loop....
netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
library (ncdf4)
for (i in 1:no_netcdf){  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working on
  brkyr <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2) # extracting the second part of the netcdf filename (which is the month)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:length(obssp)) {  
      print(j)
      de <- obssp$depth[[j]]  # a variable for the observation depth layer
      yr <- obssp$year[j]  # a variable for the observation year
      mth <- obssp$month[j]  # a variable for the observation month
      #I think i need an if statement here... if brkyr == yr & brkmth == mth do the data extraction below
        if (brkyr == yr & brkmth == mth){
          obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp) 
          obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)
        }
    }
}

str(t)
```

not extracting the correct values...

@@

shawn suggests creating a table/list of raster bricks with year and month data. Actually need to do this with the netCDFs (as you want to load the netCDF to convert)
```{r}
#the loop....
netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE)
nctable <-as.data.frame(netcdf_list)
no_netcdf <- length(netcdf_list) 
netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) 
library (ncdf4)
for (i in 1:no_netcdf){  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working on
  nctable$year[i] <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1)
  nctable$month[i] <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2)
}

```

a new loop... no this isnt going to work...
```{r warning=FALSE}
#the loop....
library (ncdf4)
netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
for (j in 1:length(obssp)) {  
      print(j)
      de <- obssp$depth[[j]]  # a variable for the observation depth layer
      yr <- obssp$year[j]  # a variable for the observation year
      mth <- obssp$month[j]  # a variable for the observation month
      if (yr == nctable$year & mth == nctable$month) {
        temp_brick <- brick(netcdf_list[i], lvar = 4)
        temp_brick <- projectRaster(temp_brick, aea) 
        for (j in 1:length(obssp)) {  
        print(j)
        de <- obssp$depth[[j]]  # a variable for the observation depth layer
        obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp) 
        obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)
        }
      }
}
```

hmmm play...
```{r}
netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
library (ncdf4)
for (i in 1:no_netcdf){  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working on
  brkyr <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2) # extracting the second part of the netcdf filename (which is the month)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:length(obssp)) {  
      print(j)
      de <- obssp$depth[[j]]  # a variable for the observation depth layer
      yr <- obssp$year[j]  # a variable for the observation year
      mth <- obssp$month[j]  # a variable for the observation month
      #I think i need an if statement here... if brkyr == yr & brkmth == mth do the data extraction below
        for (k in 1:length(obssp)){
          if (brkyr == yr & brkmth == mth){
            obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp) 
            obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)
          }
        } 
      }
}
obssp$salinity_depth
```

netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
for (i in 1:no_netcdf){  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working on
  brkyr <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2) # extracting the second part of the netcdf filename (which is the month)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:length(obssp)) {  
      print(j)
      de <- obssp$depth[[j]]  # a variable for the observation depth layer
      yr <- obssp$year[j]  # a variable for the observation year
      mth <- obssp$month[j]  # a variable for the observation month
      #I think i need an if statement here... if brkyr == yr & brkmth == mth do the data extraction below
        if (brkyr == yr & brkmth == mth){
          obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp) 
          obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)
        }
    }
}
obssp$salinity_depth
obssp$salinity_surface

### 2018-11-07 src/timesliced_random_cell_selection.Rmd
lat lon on background points done (+ little progress on the raster brick loop)

```{r}
library(sp)
xy <- background_all[ ,c(1,2)] #lon lat in col 1 and 2
sp_background <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
head(sp_background)
```
```{r}
plot(sp_background)
```

2) reproject to a wgs

@@

```{r}
head(sp_background)
```

save as csv

>>>>

```{r}
library(sp)
background_all <- as.data.frame(background_all)
xy <- background_all[ ,c(1,2)] #lon lat in col 1 and 2
sp_background <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
head(sp_background)
```

```{r}
plot(sp_background)
```

2) reproject to a wgs (use an existing file as a mask)

@@

```{r}
sp_background <- data.frame(test, long=coordinates(sp_background)[,1], lat=coordinates(sp_background)[,2])
head(sp_background)
```

save as csv
```{r}
write.csv(adddata_aea, file = "../output/bio/data_aea.csv", row.names = FALSE)
